---
title: "Greta and Causact Exploration"
subtitle: "Multi-Level Models"
author: "Peter Rabinovitch"
date: "`r Sys.time()`"
output: github_document
always_allow_html: true
---


# Introduction

The purpose of this document is to illustrate how to use the *causact* library to build some simple models. Causact allows you to create Bayesian models, and uses the *greta* library to do the computations. Well, not really - greta just sets them up, and uses TensorFlow underneath.

There are already some good resources for both [greta](https://greta-stats.org/index.html) & [causact](https://www.causact.com/), but I wanted to try some simple examples, and figured others might find this useful too.

To get everything set up, you can do no better than to follow [chapter 15 of "A Business Analyst’s Introduction to Business Analytics"](https://www.causact.com/install-tensorflow-greta-and-causact.html#install-tensorflow-greta-and-causact)

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tictoc)
library(lubridate)
library(skimr)
library(patchwork)
library(ggridges)
library(knitr)
library(DT)
library(janitor)
library(broom)
library(tictoc)
library(lme4)
library(bayesplot)
library(greta)
library(causact)

set.seed(2021)

g_fignum <- 0
figcap <- function() {
  g_fignum <<- g_fignum + 1
  return(str_c("Figure ", g_fignum))
}

g_tabnum <- 0
tabcap <- function() {
  g_tabnum <<- g_tabnum + 1
  return(str_c("Table ", g_tabnum))
}
```

Note that this document will not discuss all the things you need to do to perform a proper Bayesian analysis, that is discussed elsewhere (see the references). The purpose here is simply to show how to use causact & greta.

# Data

In order to build some simple examples we will work with one data set.

```{r, message=FALSE, warning=FALSE}
subjectdf <- tribble(
  ~condition, ~covariate, ~subject,
  "treatment", "b", "anbkv",
  "control", "b", "avpov",
  "control", "b", "ayyxo",
  "treatment", "a", "bkydk",
  "treatment", "b", "brcyb",
  "treatment", "b", "dcrhb",
  "treatment", "a", "efsoy",
  "treatment", "b", "emjwt",
  "treatment", "b", "ewlyd",
  "control", "a", "fryng",
  "treatment", "b", "fygyl",
  "control", "a", "fzvmk",
  "treatment", "a", "gjleq",
  "control", "b", "gnwyo",
  "treatment", "a", "ifxpf",
  "control", "a", "ihawn",
  "treatment", "a", "jzmgc",
  "control", "a", "lmhsz",
  "control", "b", "lusjq",
  "treatment", "a", "mbgbb",
  "treatment", "a", "nglsq",
  "treatment", "a", "ofrrl",
  "control", "a", "okyrj",
  "treatment", "b", "orbnj",
  "control", "b", "sqgvz",
  "treatment", "b", "uxabv",
  "control", "a", "xqsts",
  "control", "b", "xywyv"
)

resultsdf <- tribble(
  ~time, ~anbkv, ~avpov, ~ayyxo, ~bkydk, ~brcyb, ~dcrhb, ~efsoy, ~emjwt, ~ewlyd, ~fryng, ~fygyl, ~fzvmk, ~gjleq, ~gnwyo, ~ifxpf, ~ihawn, ~jzmgc, ~lmhsz, ~lusjq, ~mbgbb, ~nglsq, ~ofrrl, ~okyrj, ~orbnj, ~sqgvz, ~uxabv, ~xqsts, ~xywyv,
  1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
  2L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
  3L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
  4L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
  5L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
  6L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
  7L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L,
  8L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 3L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L,
  9L, 2L, 1L, 0L, 0L, 2L, 2L, 2L, 2L, 0L, 3L, 2L, 3L, 0L, 2L, 0L, 0L, 0L, 0L, 2L, 2L, 0L, 1L, 2L, 0L, 0L, 1L, 0L, 0L,
  10L, 3L, 3L, 0L, 2L, 3L, 2L, 2L, 2L, 2L, 3L, 3L, 3L, 2L, 3L, 2L, 1L, 0L, 0L, 3L, 3L, 2L, 2L, 3L, 0L, 2L, 3L, 1L, 1L,
  11L, 3L, 3L, 1L, 2L, 4L, 3L, 3L, 2L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 1L, 0L, 3L, 3L, 3L, 3L, 3L, 2L, 2L, 3L, 2L, 2L,
  12L, 3L, 3L, 3L, 3L, 5L, 3L, 3L, 3L, 3L, 3L, 3L, 2L, 3L, 3L, 3L, 3L, 3L, 0L, 3L, 3L, 3L, 3L, 3L, 3L, 2L, 3L, 2L, 3L,
  13L, 4L, 3L, 3L, 3L, 5L, 2L, 4L, 3L, 4L, 3L, 4L, 2L, 3L, 3L, 3L, 3L, 3L, 1L, 3L, 3L, 2L, 2L, 3L, 3L, 2L, 3L, 3L, 2L,
  14L, 4L, 3L, 3L, 3L, 5L, 2L, 4L, 3L, 3L, 2L, 3L, 2L, 3L, 3L, 3L, 3L, 3L, 2L, 3L, 3L, 2L, 2L, 3L, 3L, 2L, 4L, 3L, 2L,
  15L, 4L, 3L, 3L, 3L, 5L, 2L, 4L, 3L, 3L, 3L, 3L, 2L, 3L, 2L, 3L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 3L, 2L, 4L, 3L, 2L,
  16L, 5L, 2L, 3L, 3L, 5L, 2L, 4L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 3L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 3L, 2L,
  17L, 5L, 2L, 2L, 3L, 5L, 2L, 4L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L,
  18L, 5L, 2L, 2L, 3L, 5L, 2L, 4L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L,
  19L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L,
  20L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 1L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L,
  21L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L,
  22L, 5L, 1L, 2L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 1L, 2L, 2L, 2L, 2L, 3L, 1L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L,
  23L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 2L, 2L,
  24L, 5L, 2L, 3L, 2L, 5L, 2L, 5L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 3L, 2L, 2L,
  25L, 5L, 2L, 3L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L,
  26L, 5L, 2L, 3L, 2L, 5L, 2L, 5L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L,
  27L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L,
  28L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 3L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L,
  29L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L,
  30L, 5L, 2L, 2L, 2L, 5L, 2L, 5L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 2L, 3L, 0L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L
)

resultsdfl <- resultsdf %>%
  pivot_longer(
    cols = -time,
    names_to = "subject",
    values_to = "outcome"
  )

df <- resultsdfl %>%
  left_join(subjectdf, by = c("subject" = "subject")) %>%
  arrange(subject, time)

rm(resultsdfl, resultsdf, subjectdf)
```

# EDA

Ok, so what does this data set look like?


# Simple Models


## Model: lm(outcome~1)


## Model: lm(outcome~covariate+condition)




## Model: lm(outcome~covariate+condition+subject) with plates


```{r, message=FALSE, warning=FALSE, fig.height=8, fig.cap=figcap()}
graph <- dag_create() %>%
  dag_node("outcome", "o", rhs = normal(mu, 3), data = df$outcome) %>%
  dag_node("mu", "mu", rhs = condeffect + coveffect + timeeffect + subeffect, child = "o") %>%
  dag_node("condeffect", "condeffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_node("coveffect", "coveffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_node("timeeffect", "timeeffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_node("subeffect", "subeffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_plate("Condition Effect", "i", nodeLabels = c("condeffect"), data = df$condition, addDataNode = TRUE) %>%
  dag_plate("Covariate Effect", "j", nodeLabels = c("coveffect"), data = df$covariate, addDataNode = TRUE) %>%
  dag_plate("Time Effect", "t", nodeLabels = c("timeeffect"), data = df$timec, addDataNode = TRUE) %>%
  dag_plate("Subject Effect", "s", nodeLabels = c("subeffect"), data = df$subject, addDataNode = TRUE) %>%
  dag_plate("Observation", "b", nodeLabels = c("o"))
graph %>% dag_render()
```
```{r, message=FALSE, warning=FALSE, fig.cap=figcap()}
drawsDF <- graph %>% dag_greta()
drawsDF %>% dagp_plot()
```

## Model: lm(outcome~covariate+condition+timec+subject) 

This is just like model 6, but we use the plate notation for the outcome.

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.cap=figcap()}
graph <- dag_create() %>%
  dag_node("outcome", "o", rhs = normal(mu, 3), data = df$outcome) %>%
  dag_node("mu", "mu", rhs = condeffect + coveffect + timeeffect + subeffect, child = "o") %>%
  dag_node("condeffect", "condeffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_node("coveffect", "coveffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_node("timeeffect", "timeeffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_node("subeffect", "subeffect", rhs = uniform(0, 5), child = "mu") %>%
  dag_plate("Condition Effect", "i", nodeLabels = c("condeffect"), data = df$condition, addDataNode = TRUE) %>%
  dag_plate("Covariate Effect", "j", nodeLabels = c("coveffect"), data = df$covariate, addDataNode = TRUE) %>%
  dag_plate("Time Effect", "t", nodeLabels = c("timeeffect"), data = df$timec, addDataNode = TRUE) %>%
  dag_plate("Subject Effect", "s", nodeLabels = c("subeffect"), data = df$subject, addDataNode = TRUE) %>%
  dag_plate("Observation", "b", nodeLabels = c("o"))
graph %>% dag_render()
```
```{r, message=FALSE, warning=FALSE, fig.cap=figcap()}
drawsDF <- graph %>% dag_greta()
drawsDF %>% dagp_plot()
```




# To Do


# Conclusion


# Appendices

<details>

<summary>

References

</summary>

[greta](https://greta-stats.org/index.html)  
[causact](https://www.causact.com/)  
[Stan](https://mc-stan.org/)  
[Rethinking](https://xcelab.net/rm/statistical-rethinking/)  
[RAOS](https://avehtari.github.io/ROS-Examples/index.html)  
[Workflow 1](http://www.stat.columbia.edu/~gelman/research/unpublished/Bayesian_Workflow_article.pdf)  
[Workflow 2](https://betanalpha.github.io/assets/case_studies/principled_bayesian_workflow.html)  

</details>

<details>

<summary>

SessionInfo

</summary>

```{r}
sessionInfo()
```

</details>
